A sheaf--function dictionary for algebraic tori over local fields

### 2013-08-20 shanghai.tex
I removed the proposition on weil restriction from the final section. The paper still works!

\begin{proof}
Observe that $\QC(\GN{T})$ comes equipped with monoidal subcategories
\[
\begin{tikzcd}
\ & & & \QC(\GN{T})\\
\arrow{urrr}{F_n} \QC(\GN{T}_n) \arrow[swap]{rr}{f_{n\leq m}^*} && \QC(\GN{T}_m) \arrow[swap]{ur}{F_m} & \\
\end{tikzcd}
\]
defined by $F_n(\qcs{L}) = [\qcs{L}_n]$ and $F_n(\alpha_n) = [\alpha_n]$, for all positive integers $n\leq m$.
Passing to isomorphism classes of objects and invoking the universal property of $\varinjlim_{d\in \ZZ_{>0}} \QCiso{\GN{T}_d}$ gives a group homomorphism
$\theta: \varinjlim_{d\in \ZZ_{>0}} \QCiso{\GN{T}_d} \to \QCiso{\GN{T}}$; this is a bijection 
with inverse $[\qcs{L}]_{/\text{iso}} \mapsto [\qcs{L}_{/\text{iso}}]$.
Now use Theorem~\ref{thm:application}.
\end{proof}

@@

As indicated in the Introduction, we are tantalized 
by the idea that quasicharacter sheaves for $T$ 
may play a key role in a new geometric, categorial version 
of the local Langlands correspondence for algebraic tori 
over non-archimedean local fields. 
As with the local Langlands correspondence for algebraic tori over local fields, 
it should be possible to reduce the essential points of the argument to the case $T= \Gm{K}$,
provided that the category of quasicharacter sheaves behaves properly with regard to
Weil restriction; see \cite{yu:09a}*{Thm 7.5 (2)}. 
The next proposition shows that the category of quasicharacter sheaves for 
algebraic tori does indeed behave properly with regard to Weil restriction.

\begin{proposition}
Let $K'/K$ be a finite Galois extension; 
let $k'/k$ be the corresponding finite Galois extension of residue fields.
Let $T$ be an algebraic torus over $K$ and 
set $T' = T \times_\Spec{K} \Spec{K'}$.  
The natural inclusion $T \hookrightarrow \Res_{K'/K} T'$ 
induces a map of $k$-schemes
\[
\GN{T} \hookrightarrow \Res_{k'/k} \GN{T}'
\]
and thus a functor on quasicharacter sheaves through pullback.
\[
\QC(\Res_{k'/k} \GN{T}') \to \QC(\GN{T}).
\]
Through trace of Frobenius, this functor induces the homomorphism
\[
\begin{tikzcd}[column sep=60]
\Hom_{\text{qc}}(T(K'), \EEx) \arrow{r}{\chi \mapsto \chi\vert_{T(K)}} &\Hom_{\text{qc}}(T(K), \EEx).
\end{tikzcd}
\] 
\end{proposition}

\begin{proof}
Weil restriction behaves properly with respect to formation of Néron models:
adapting the argument in \cite{bosch-lutkebohmert-reynaud:NeronModels}*{\S 7.6, Prop 6} 
gives a canonical isomorphism 
\Clifton{I need to check this.}
\begin{equation}
\underline{\Res_{K'/K}(T')} \iso \Res_{\mathcal{O}_{K'}/\OK}(\underline{T'}).
\end{equation}
And Weil restriction behaves properly with respect to the Greenberg transform:
\Clifton{I'm not so sure about this . . . In fact, I don't believe it!}
\begin{equation}
\Gr_{n}\left( \Res_{\mathcal{O}_{K'}/\OK}(\underline{T'})\right)
\iso
\Res_{k'/k} \Gr_n(\underline{T'}),
\end{equation}
where $k'$ is the residue field for $K'$.
And the Greenberg transform respects products:
\[
\Gr_n(\underline{T'}) = \Gr_n(\underline{T}\times_{k} {k'}) \iso \Gr_n(\underline{T})\times_{k} {k'} = \Gr_n(\underline{T})_{k'}.
\]
Thus, the 
Now use Proposition~\ref{prop:csbe}.
\end{proof}

This paper has shown how quasicharacter sheaves 
on the Greenberg transform of the N\/eron model of $T$ 
determine admissible characters of $T(K)$, and vice versa, functorially.
When coupled with the reciprocity map $\text{rec}_T $ for $T$, 
which enjoys the same functorial properties, 
it follows that quasicharacter sheaves for $T$ 
determine Langlands parameters for $T$, and vice versa, functorially.
\[
\begin{tikzcd}
\ & \arrow[swap]{dl}{\trFrob{G}} \QCiso{\GN{T}} \arrow[dashed]{dr}  & \\
\Hom_{\text{qc}}(T(K),\EEx) \arrow{rr}{\text{rec}_T} && H^1(K,\hat{T}_\ell) \\
\end{tikzcd}
\]
We suspect that it is possible to extract Langlands parameters 
from quasicharacter sheaves directly and that the reciprocity map 
can then be be viewed as a matching between quasicharacters and Langlands parameters coming
 from the same quasicharacter sheaf. 
 Results in this paper provide some tools for reduction of the problem to the case $T=\Gm{K}$. 
 In that case, our suspicions are all but confirmed 
 by the class field theory of Serre-Hazewinkel
 \citelist{\cite{serre:60a},\cite{serre:61a},\cite{demazure-gabriel:GroupesAlgebriques}*{Appendice}},
 especially as revisited in \cite{suzuki-yoshida:12a}*{\S~2}. 
 What remains to be done is a careful comparison of 
 quasicharacter sheaves for $\Gm{K}$ and certain local systems on
  the fpqc-site of the pro-quasi-algebraic varieties appearing in \cite{suzuki-yoshida:12a}.

>>>>

\begin{proof}
Observe that $\QC(\GN{T})$ comes equipped with monoidal subcategories
\[
\begin{tikzcd}
\ & & & \QC(\GN{T})\\
\arrow{urrr}{F_n} \QC(\GN{T}_n) \arrow[swap]{rr}{f_{n\leq m}^*} && \QC(\GN{T}_m) \arrow[swap]{ur}{F_m} & 
\end{tikzcd}
\]
defined by $F_n(\qcs{L}) = [\qcs{L}_n]$ and $F_n(\alpha_n) = [\alpha_n]$, for all positive integers $n\leq m$.
Passing to isomorphism classes of objects and invoking the universal property of $\varinjlim_{d\in \ZZ_{>0}} \QCiso{\GN{T}_d}$ gives a group homomorphism
$\theta: \varinjlim_{d\in \ZZ_{>0}} \QCiso{\GN{T}_d} \to \QCiso{\GN{T}}$; this is a bijection 
with inverse $[\qcs{L}]_{/\text{iso}} \mapsto [\qcs{L}_{/\text{iso}}]$.
Now use Theorem~\ref{thm:application}.
\end{proof}

@@

As indicated in the Introduction, we are tantalized 
by the idea that quasicharacter sheaves for $T$ 
may play a key role in a new geometric, categorial version 
of the local Langlands correspondence for algebraic tori 
over non-archimedean local fields. 
%
This paper has shown how quasicharacter sheaves 
on the Greenberg transform of the N\/eron model of $T$ 
determine admissible characters of $T(K)$, and vice versa, functorially.
When coupled with the reciprocity map $\text{rec}_T $ for $T$, 
which enjoys the same functorial properties, 
it follows that quasicharacter sheaves for $T$ 
determine Langlands parameters for $T$, and vice versa, functorially.
\[
\begin{tikzcd}
\ & \arrow[swap]{dl}{\trFrob{G}} \QCiso{\GN{T}} \arrow[dashed]{dr}  & \\
\Hom_{\text{qc}}(T(K),\EEx) \arrow{rr}{\text{rec}_T} && H^1(K,\hat{T}_\ell) 
\end{tikzcd}
\]
We suspect that it is possible to extract Langlands parameters 
from quasicharacter sheaves directly and that the reciprocity map 
can then be be viewed as a matching between quasicharacters and Langlands parameters coming
 from the same quasicharacter sheaf. 
 Results in this paper provide some tools for reduction of the problem to the case $T=\Gm{K}$. 
 In that case, our suspicions are all but confirmed 
 by the class field theory of Serre-Hazewinkel
 \citelist{\cite{serre:60a},\cite{serre:61a},\cite{demazure-gabriel:GroupesAlgebriques}*{Appendice}},
 especially as revisited in \cite{suzuki-yoshida:12a}*{\S~2}. 
 What remains to be done is a careful comparison of 
 quasicharacter sheaves for $\Gm{K}$ and certain local systems on
  the fpqc-site of the pro-quasi-algebraic varieties appearing in \cite{suzuki-yoshida:12a}.
